\clearpage
% \setcounter{page}{1}
% \maketitlesupplementary

\begin{appendices}
\section{Structured Captions}
\label{app:structured-json}

In this section we describe our structured captioning strategy. 
Each caption is a long, detailed JSON that decomposes the image into interpretable fields, covering the objects, background, lighting, aesthetics, photographic attributes, style, and text. This structure guarantees full semantic coverage and improves controllability during training and inference.
The structure of our JSON is described in Table~\ref{tab:JSON-structure}.\input{tables/JSON_structure} 

Our structured captions are generated using Gemini 2.5 with a dedicated system prompt (see Table~\ref{tab:system_prompt}) and using Gemini's structured outputs configured with our JSON schema (see Table~\ref{tab:strcutured_outputs}) that enforce completeness, consistency, and adherence to the predefined schema. 
\input{tables/JSON_system_prompt}
\input{tables/JSON_gemini_strcture_output}





\section{Additional Training Details}\label{sec:additional-training-details}

The data distribution is illustrated in Figure~\ref{fig:data_distribution}.
\input{Figures/data_distribution/data_distribution}


We train the model using the AdamW optimizer~\cite{loshchilov2017decoupled} with weight decay of $1\times10^{-4}$, $\beta_1=0.9$, $\beta_2=0.999$, and $\epsilon=1\times10^{-15}$. The learning rate is set to $1\times10^{-4}$ with a constant schedule and a warmup of $10$K steps, followed by an additional $2.5$K warmup steps when increasing resolution. Training follows the flow-matching formulation~\cite{lipman2023flowmatchinggenerativemodeling}, with a logit-normal noise schedule combined with resolution-dependent timestep shifting~\cite{esser2024scaling}.


We train the model in stages of increasing resolution. The first stage is conducted at $256^2$ resolution for $652$k steps, where the initial $300$k steps employ REPA~\cite{yu2024representation} using DINOv2-L encodings~\cite{oquab2023dinov2} and REPA coefficient of 0.1. Training then continues for an additional $100$k steps at $512^2$ resolution, followed by $50$k steps at $1024^2$. At each resolution stage, the effective batch size starts at approximately $1$K and is gradually increased to $2$K as training progresses.

Post-training, we perform aesthetic finetuning with 3,000 hand-picked images, followed by DPO training~\cite{wallace2024diffusion} with dynamic beta~\cite{liu2025improving} to improve text rendering.





\section{Additional Ablation Studies Details}
\label{sec:appendix_ablation}

We provide implementation details for the ablation studies discussed in Section~\ref{sec:captions} (long structured captions vs. short captions) and Section~\ref{sec:dimfusion} (convergence behavior across architectures).
To reduce compute cost, all ablations were conducted under identical training and sampling conditions using 1B-parameter models trained with the SDXL VAE \cite{podell2023sdxl}.
Table \ref{tab:ablations_model_arch} lists the architectural hyperparameters used in all experiments.
\input{tables/ablation_studies_arch}

In Section \ref{sec:captions}, we examine the effect of caption length using the \emph{DimFusion} architecture with \emph{SmolLM3-3B} as the LLM, trained once with long structured captions and once with short captions.
In Section \ref{sec:dimfusion}, we compare three architectures trained on long structured captions.
The baseline follows \cite{esser2024scaling} and uses T5 \cite{raffel2020exploring} as the text encoder.
We further evaluate two large-language-model fusion variants that share the same hyperparameters as the baseline:
(1) \emph{DimFusion} with \emph{SmolLM3-3B} as the LLM; and
(2) \emph{TokenFusion}, following HiDream \cite{cai2025hidream}, concatenates SmolLM3-3B hidden states with the text embeddings along the \emph{sequence} dimension, effectively doubling the number of text tokens in each attention layer.

All models were trained on 70M licensed image-caption pairs, with the same optimizer, learning rate and noise scheduling described in Appendix~\ref{sec:additional-training-details}, and effective batch size of 1024. We also employed REPA~\cite{yu2024representation} with DINOv2-L encodings~\cite{oquab2023dinov2} and a REPA coefficient of 0.5.
For evaluation, FID was computed on 30K samples from the COCO-2014~\cite{lin2014microsoft} validation split, using 50 inference steps and no classifier-free guidance.


\section{Additional Samples from \modelname{}}

Additional samples from \modelname{} in various aspect ratios appear in Figure~\ref{fig:app_samples}, and additional examples for refinement and disentanglement appear in Figure~\ref{fig:app_disentanglement}.
\input{Figures/samples/app_samples}
\input{Figures/disentanglement/app_dis_fig}


% \section{User Study}
% Print screens to demonstrate the user interface for our user study are provided in Figure~\ref{fig:user_study_ui}.

% \input{Figures/user_study/user_study}

\end{appendices}

